Automatic Identification and Annotation of Parallel Tasks in Structured Programs

### 2017-11-17 docs/ACM/paper.tex
Adding an abstract to the paper.

%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

@@

\begin{abstract}
We need to write our abstract here.
\end{abstract}

@@

Annotation systems have risen to a place of prominence as a simple and
effective way to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11}, Cilk++~\cite{Leiserson09} and
Tareador~\cite{Ayguade15}.
Annotations work as a meta-language: they let developers grant parallel
semantics to syntax originally written to execute sequentially.
The compiler-related literature contains a generous load of success stories
describing the use of annotation systems to parallelize programs.
Combined with modern accelerators, they have led to substantial performance
gains~\cite{Bertolli14,Mendonca17,Poesia17,Reyes12,Wienke12}.
Yet, even if convenient, the use of annotations is not straightforward, and
lacks supporting tools.

Annotations such as OpenMP or OpenACC still require developers to
worry about typical hassles of the parallel world, such as race conditions
or deadlocks.
Moreover, because they are used in tandem with imperative programming languages,
such troubles are only made worse by pointer aliasing.
There exist tools that insert automatic annotations in
programs~\cite{Mendonca16,Pingali11,Nugteren14}.
All these technologies explore data-parallelism -- the possibility of running
the same computation independently on different data.
Task parallelism remains still uncharted land, in what concerns automatic
annotation.
Such fact is unfortunate, because much of the power of current annotation
systems lays on their ability to create tasks~\cite{Ayguade09}.
This ability brings said systems closer to irregular programs such as those
that implement graphs and worklist algorithms \`{a} la
Pingali~\cite{Ben-Nun17,Kulkarni11,Pingali11}.
The goal of this work is to address this omission.

This paper describes \Taskminer, a compiler that mines task parallelism
from programs.
To fulfill this goal \Taskminer{} solves many challenges.
First, it finds program regions, e.g., loops or functions, that can be
effectively mapped onto tasks (Section~\ref{sub:identification}).
Second, it determines symbolic bounds to all the memory blocks accessed within
those regions (Section~\ref{sub:symb}).
Third, it maps such information back into source code, producing annotations
that programmers can read (Section~\ref{sub:ir}).
Fourth, it extracts parameters from the code to estimate the profitability of
tasks, given the cost to create them.
These parameters feed conditional checks, which, at runtime, enable or disable
the creation of tasks (Section~\ref{sub:profit}).
Finally, \Taskminer{} determines which program variables need to be replicated at
each new task, or shared among them (Section~\ref{sub:variance}).
This paper describes the design and the implementation of these technologies, and
demonstrate their benefit.

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.95\columnwidth]{images/ex_Regions}
\caption{Write caption here.}
\label{fig:ex_Regions}
\end{center}
\end{figure}

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.95\columnwidth]{images/ex_cutoff}
\caption{Write caption here.}
\label{fig:ex_cutoff}
\end{center}
\end{figure}

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\columnwidth]{images/ex_privatize}
\caption{Write caption here.}
\label{fig:ex_privatize}
\end{center}
\end{figure}

@@

We aimed to show that our tool exceeded in three main aspects that surround parallel programming. Thus our work evaluation stands on these 3 pillars:

@@

\begin{itemize}

\item Practicality
\item Applicability
\item Performance

\end{itemize}

\textbf{\textit{Practicality}}. Programming in a parallel pattern is deemed difficult by programmers mainly because of the complexity of its data structures and algorithms. We show that our tool is simple and practical, for it does not involve changing the code and data structures in order to fit in a parallel pattern. The code is simply annotated by compiler directives.

\textbf{\textit{Applicability}}.

\textbf{\textit{Performance}}. When it comes to parallel computing, it is undeniable that the main achievable goal should be performance. There's no reason why programmers would opt for parallelism other than the desire for faster programs. Therefore, we ought to show that the programs annotated with TaskMiner are faster than their serial counterparts.

>>>>

%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{IEEE}{2018}{Planet Earh}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

@@

\begin{abstract}
This paper reports the design and implementation of a suit of static analysis
and code generation techniques to annotate programs with OpenMP pragmas for
task parallelism.
These techniques deal with problems such as the discovery and use of program
symbols to estimate the profit of tasks, bound their recursive depth and limit
the memory regions onto which they operate.
These techniques have been implemented into the first source-to-source
compiler able to insert OpenMP pragmas into C/C++ programs without human
intervention.
By building onto the solid collection of static analyses available in LLVM,
and relying on OpenMP's runtime ability to disambiguate pointers, we show that
we can annotate large and convoluted programs, often replicating the performance
gains of handmade annotation.
\end{abstract}

@@

Annotation systems have risen to a place of prominence as a simple and
effective way to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11}, Cilk++~\cite{Leiserson09} and
Tareador~\cite{Ayguade15}.
Annotations work as a meta-language: they let developers grant parallel
semantics to syntax originally written to execute sequentially.
%The compiler-related literature contains a generous load of success stories
%describing the use of annotation systems to parallelize programs.
Combined with modern accelerators, they have led to substantial performance
gains~\cite{Bertolli14,Mendonca17,Poesia17,Reyes12,Wienke12}.
Yet, even if convenient, the use of annotations is not straightforward, and
lacks supporting tools.

Annotations such as OpenMP or OpenACC still require developers to
worry about typical hassles of the parallel world, such as race conditions
or deadlocks.
Moreover, because they are used in tandem with imperative programming languages,
such troubles are only made worse by pointer aliasing.
There exist tools that insert automatic annotations in
programs~\cite{Amini12,Guelton12,Mendonca16,Pingali11,Nugteren14}.
All these technologies explore data-parallelism -- the possibility of running
the same computation independently on different data.
Task parallelism remains still uncharted land, in what concerns automatic
annotation.
Such fact is unfortunate, because much of the power of current annotation
systems lays on their ability to create tasks~\cite{Ayguade09}.
This ability brings said systems closer to irregular programs such as those
that implement graphs and worklist algorithms \`{a} la
Pingali~\cite{Ben-Nun17,Kulkarni11,Pingali11}.
The goal of this work is to address this omission.

This paper describes \Taskminer, a compiler that mines task parallelism
from programs.
To fulfill this goal \Taskminer{} solves many challenges.
First, it finds program regions, e.g., loops or functions, that can be
effectively mapped onto tasks (Section~\ref{sub:identification}).
Second, it determines symbolic bounds to all the memory blocks accessed within
those regions (Section~\ref{sub:symb}).
Third, it extracts parameters from the code to estimate the profitability of
tasks, given the cost to create them.
These parameters feed conditional checks, which, at runtime, enable or disable
the creation of tasks (Section~\ref{sub:profit}) and limit their recursion
depth (Section~\ref{sub:rec}).
Additionally, \Taskminer{} determines which program variables need to be replicated at
each new task, or shared among them (Section~\ref{sub:variance}).
Finally, it maps all this information back into source code, producing annotations
that programmers can read (Section~\ref{sub:ir}).
This paper describes the design and the implementation of these technologies, and
demonstrate their benefit.

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.95\columnwidth]{images/ex_Regions}
\caption{Challenges~\ref{ch:Regions} and~\ref{ch:cost}: identifying memory regions
with symbolic limits, and using runtime information to estimate the profit of
tasks.}
\label{fig:ex_Regions}
\end{center}
\end{figure}

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=0.95\columnwidth]{images/ex_cutoff}
\caption{Challenge~\ref{ch:cutoff}: bounding the creation of recursive tasks with
counter associated with function calls.}
\label{fig:ex_cutoff}
\end{center}
\end{figure}

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\columnwidth]{images/ex_privatize}
\caption{Challenge~\ref{ch:privatize}: variable \textsf{j} must be replicated among
tasks, to avoid the occurrence of data races.}
\label{fig:ex_privatize}
\end{center}
\end{figure}

@@

%We aimed to show that our tool exceeded in three main aspects that surround parallel programming. Thus our work evaluation stands on these 3 pillars:

@@

%\begin{itemize}

%\item Practicality
%\item Applicability
%\item Performance

%\end{itemize}

%\textbf{\textit{Practicality}}. Programming in a parallel pattern is deemed difficult by programmers mainly because of the complexity of its data structures and algorithms. We show that our tool is simple and practical, for it does not involve changing the code and data structures in order to fit in a parallel pattern. The code is simply annotated by compiler directives.

%\textbf{\textit{Applicability}}.

%\textbf{\textit{Performance}}. When it comes to parallel computing, it is undeniable that the main achievable goal should be performance. There's no reason why programmers would opt for parallelism other than the desire for faster programs. Therefore, we ought to show that the programs annotated with TaskMiner are faster than their serial counterparts.
